---
title: "Loading data into R from Network Canvas - Networks 2021"
output:
  html_notebook: default
  html_document:
    df_print: paged
---

This Notebook provides basic input and example analysis of data exported from the [*Network Canvas*](http://networkcanvas.com) software. The example will use simulated data from the following network canvas [protocol file](need to add github link to protocol file) which can be opened in the Network Canvas Interviewer or Architect apps.

### Import and clean the data 

The first step is to load in data from the github repository. We use the *egor* package to load in the data, using examples from an egor [vignette](https://cran.r-project.org/web/packages/egor/vignettes/using_egor.html)


```{r,message = FALSE}
# Load egor
library(egor)
library(sna)

# Specify your folder path here 
folderPath <- "/Folder/Path/"

# Read each type of file into list, combine into single data frame
alterData <- folderPath %>%
  list.files(full.names=TRUE,pattern="attributeList_Person.csv") %>%
  lapply(read.csv) %>%
  bind_rows()

edgelistData <- folderPath %>%
  list.files(full.names=TRUE,pattern="edgeList_friends.csv") %>%
  lapply(read.csv) %>%
  bind_rows()

egoData <- folderPath %>%
  list.files(full.names=TRUE,pattern="ego.csv") %>%
  lapply(read.csv) %>%
  bind_rows()
```

Now we can examine if the data looks like it should from the alter file. As you'll see, the alter data file contains a unique alter identifier (i.e., "networkCanvasUUID") as well as a unique ego identifier (i.e., "networkCanvasEgoUUID") which will both be used to define the egor object.
```{r}
head(alterData)
```

We'll want to do a little data cleaning by recoding family member status and communication frequency, before loading it into an egor object.
```{r}
alterData$family_member2 <- alterData$family_member
alterData$family_member2[is.na(alterData$family_member2) | alterData$family_member2==""] <- "false"

alterData$communication_freq2 <- alterData$communication_freq
alterData$communication_freq2[alterData$communication_freq2==-1] <- "Less than \n monthly"
alterData$communication_freq2[alterData$communication_freq2==1] <- "Monthly"
alterData$communication_freq2[alterData$communication_freq2==2] <- "Weekly"
alterData$communication_freq2[alterData$communication_freq2==3] <- "Daily"

```

After loading and cleaning the data, we still need to make it into an egor object
```{r}
# Load the file into R
egorNetworkCanvas <- egor(alters = alterData,
           egos = egoData,
           aaties = edgelistData,
           ID.vars = list(
             ego = "networkCanvasEgoUUID",
             alter = "networkCanvasUUID",
             source = "networkCanvasSourceUUID",
             target = "networkCanvasTargetUUID"))
```

### Data visualization

Lets start with a simple visualization of one ego network. To do this we'll first convert it to a 'network' object and use the gplot function from the sna package. This visualization shows the new node labels and colors each node by the communication frequency with the participant.
```{r}

firstEgoNet <- as_network(egorNetworkCanvas)[[2]]
firstEgoNet%v%"vertex.names" <- firstEgoNet%v%"name"

colorScheme <- c( "#CC6677", "#117733", "#AA4499", 
                  "#6699CC")

# A little recoding to get a color for each frequency
nodeColors <- ifelse(firstEgoNet%v%"communication_freq"=="-1",colorScheme[1],
                    ifelse(firstEgoNet%v%"communication_freq"=="1",colorScheme[2],
                           ifelse(firstEgoNet%v%"communication_freq"=="2",colorScheme[3],
                                  colorScheme[4])))

gplot(firstEgoNet,
       usearrows = FALSE,
       label = firstEgoNet%v%"name",
       displaylabels = TRUE,
       vertex.col=nodeColors,
       edge.col="gray")
```

As you can see, this only shows a single egocentric network. However, the egor package has several functions that facilitate comparison of networks across ego nets. For example, here is a visualization showing each ego net with nodes location being dependent on their communciation with the participant and their status as a family member (i.e., true/false).

```{r}
# Make a visualization displaying both frequency of communication and family member status
plot(egorNetworkCanvas, venn_var = "family_member2", pie_var = "communication_freq2", type = "egogram")
```

## Going to do some basic analysis here. Maybe degree distributions and a bar chart showing communication frequency 
